---
title: "Frontloading_RMD"
author: "Samuel Gurr"
date: "9/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
# LOAD PACKAGES
library(dplyr)
library(reshape2)
library(ggplot2)
library(kableExtra)
library(knitr)
# SET WORKING DIRECTORY AND LOAD DATA
setwd("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/")

 # LOAD DATA 
annot = read.delim2(file = "Seq_details/Panopea-generosa-genes-annotations.txt",header = F) # annotation file P generosa 

# WGCNA modules
day7.ModMem   <- read.csv(file="Analysis/Output/WGCNA/subseq_treatments_all/Day7/d7.WGCNA_ModulMembership.csv", sep=',', header=TRUE)   %>%  dplyr::rename(Pgen_ID = X) %>%  mutate(Pgen_ID = as.character(Pgen_ID)) %>%  dplyr::select(c('Pgen_ID','moduleColor'))
day14.ModMem  <- read.csv(file="Analysis/Output/WGCNA/subseq_treatments_all/Day14/d14.WGCNA_ModulMembership.csv", sep=',', header=TRUE)  %>%  dplyr::rename(Pgen_ID = X) %>%  mutate(Pgen_ID = as.character(Pgen_ID)) %>%  dplyr::select(c('Pgen_ID','moduleColor'))
day21.ModMem  <- read.csv(file="Analysis/Output/WGCNA/subseq_treatments_all/Day21/d21.WGCNA_ModulMembership.csv", sep=',', header=TRUE)  %>%  dplyr::rename(Pgen_ID = X) %>%  mutate(Pgen_ID = as.character(Pgen_ID)) %>%  dplyr::select(c('Pgen_ID','moduleColor'))

# expression read counts 
# Day 7
Day7_exp_data <- read.csv(file="Analysis/Output/DESeq2/10cpm/Day7/ Day7.PrimaryDEGs_AvM__DESeq2results.csv", sep=',', header=TRUE) %>% dplyr::select(!c('X','baseMean','log2FoldChange','lfcSE','stat','pvalue','padj'))
# Day 14
Day14_exp_data <- read.csv(file="Analysis/Output/DESeq2/10cpm/Day14/ Day14.Primary_DESeq2results.csv", sep=',', header=TRUE) %>% dplyr::select(!c('X','baseMean','log2FoldChange','lfcSE','stat','pvalue','padj'))
# Day 21
Day21_exp_data <- read.csv(file="Analysis/Output/DESeq2/10cpm/Day21/ Day21.Primary_DESeq2results.csv", sep=',', header=TRUE) %>% dplyr::select(!c('X','baseMean','log2FoldChange','lfcSE','stat','pvalue','padj'))

# experiment metadata (treatments)
Day7_Master.Exp.Metadata   <- read.csv(file="Analysis/Data/Experiment_Metadata/Master_Phyenotype.and.Exp.Treatment_Metadata.csv", sep=',', header=TRUE) %>% 
  dplyr::filter(Date %in% 20190731) %>% # call only day 7 
  dplyr::select(c('Sample.Name', 'All_Treatment')) %>% # select only the sample names and the treatment groups 
  dplyr::mutate(All_Treatment = factor(All_Treatment)) # make the treatment groups a factor 
Day14_Master.Exp.Metadata   <- read.csv(file="Analysis/Data/Experiment_Metadata/Master_Phyenotype.and.Exp.Treatment_Metadata.csv", sep=',', header=TRUE) %>% 
  dplyr::filter(Date %in% 20190807) %>% # call only day 7 
  dplyr::select(c('Sample.Name', 'All_Treatment')) %>% # select only the sample names and the treatment groups 
  dplyr::mutate(All_Treatment = factor(All_Treatment)) # make the treatment groups a factor 
Day21_Master.Exp.Metadata   <- read.csv(file="Analysis/Data/Experiment_Metadata/Master_Phyenotype.and.Exp.Treatment_Metadata.csv", sep=',', header=TRUE) %>% 
  dplyr::filter(Date %in% 20190814) %>% # call only day 7 
  dplyr::select(c('Sample.Name', 'All_Treatment')) %>% # select only the sample names and the treatment groups 
  dplyr::mutate(All_Treatment = factor(All_Treatment)) # make the treatment groups a factor 
```

## Call modules representing higher expression by naive clams throughout the subseqent exposures 

- 'Naive_modules' == all mods of interest

```{r naive mods, echo=TRUE}
day7.brown    <- day7.ModMem %>% dplyr::filter(moduleColor %in% 'brown')     %>% dplyr::mutate(day ='Day7')
day14.brown   <- day14.ModMem %>% dplyr::filter(moduleColor %in% 'brown')    %>% dplyr::mutate(day ='Day14')
day21.blue    <- day21.ModMem %>% dplyr::filter(moduleColor %in% 'blue')     %>% dplyr::mutate(day ='Day21')
day21.magenta <- day21.ModMem %>% dplyr::filter(moduleColor %in% 'magenta')  %>% dplyr::mutate(day ='Day21')

Naive_modules <- rbind(day7.brown, day14.brown, day21.blue, day21.magenta)

```

- 'NaiveResponse_genes_data' call all genes that occured THREE times in 'Naive_modules'
- Why? These genes are represent those with persistant high expression relative to the pre-exposed (primed) clams!
- there are 315 total genes in this category

```{r persistantly expressed by naive, echo=T}

NaiveResponse_genes <- Naive_modules %>% 
                        dplyr::group_by(Pgen_ID) %>% 
                        dplyr::summarise(count = n()) %>% 
                        dplyr::filter(count == 3)
# View(NaiveResponse_genes) 
NaiveResponse_genes        = NaiveResponse_genes$Pgen_ID
NaiveResponse_genes_ANNOT  = match(NaiveResponse_genes, annot$V1)
NaiveResponse_genes_data   <- data.frame(geneSymbol = annot$V1[NaiveResponse_genes_ANNOT],
                                         Annotation = annot$V7[NaiveResponse_genes_ANNOT])
# nrow(NaiveResponse_genes_data) # 315 genes present on all sampling days 
```

# Calculate the contorl ratio (Y axis) and the foldchangie ratio (x axis) 

```{r merge with expression data for plotting, echo=T}
# Day 7 gene expression
Day_WGCNA_genes <- Day7_exp_data %>% 
  dplyr::filter(Gene %in% NaiveResponse_genes)
Day7_WGCNA_genes_melted     <- Day_WGCNA_genes %>% 
  reshape2::melt(id.var = 'Gene') %>% 
  dplyr::rename(Sample.Name = variable)
Day7_WGCNA_genes_Merge <- merge(Day7_WGCNA_genes_melted, Day7_Master.Exp.Metadata, by = 'Sample.Name') %>% 
  dplyr::group_by(Gene, All_Treatment) %>% 
  dplyr::select(!'Sample.Name') %>% 
  dplyr::summarise(meanExp = mean(value))
Day7_READY <- dcast(Day7_WGCNA_genes_Merge, Gene ~ All_Treatment)
for (i in 1:nrow(Day7_READY)) {
          # Moderate - higher expression AM > AA
        if (Day7_READY$AM[i] > Day7_READY$AA[i]) {
          Day7_READY$wgcna.xall_mod[i] <- ( (Day7_READY$MM[i] / Day7_READY$MA[i]) / (Day7_READY$AM[i] / Day7_READY$AA[i]) ) # call MA as the control for MM ratio
          Day7_READY$wgcna.yall_mod[i] <- (Day7_READY$MA[i] / Day7_READY$AA[i]) # Y Axis - this is simply the conditioned control over the naive control ( MA / AA ) 
              } else {
                Day7_READY$wgcna.xall_mod[i] <- NA  # X axis  - call NA    
                Day7_READY$wgcna.yall_mod[i] <- NA # Y Axis  - call NA
              }
            # Severe - higher expression AS > AA
          if (Day7_READY$AS[i] > Day7_READY$AA[i]) {
            Day7_READY$wgcna.xall_sev[i] <- ( (Day7_READY$MS[i] / Day7_READY$MA[i]) / (Day7_READY$AS[i] / Day7_READY$AA[i]) ) # call AA as the control for MM ratio
            Day7_READY$wgcna.yall_sev[i] <- (Day7_READY$MA[i] / Day7_READY$AA[i])            # Y Axis - this is simply the conditioned control over the naive control ( MA / AA ) 
                } else {
                  Day7_READY$wgcna.xall_sev[i] <- NA  # X axis  - call NA 
                  Day7_READY$wgcna.yall_sev[i] <- NA # Y Axis  - call NA
                }
}
# Day7_READY
x = data.frame(Gene = annot$V1[(match(Day7_READY$Gene, annot$V1))],
               Gene_Description = annot$V7[(match(Day7_READY$Gene, annot$V1))])
Day7_Frontload   <- merge(x, Day7_READY, by = 'Gene')
Day7_Frontload_2 <- Day7_Frontload %>% 
                      dplyr::mutate(baseMeanNAIVE_control = AA) %>% 
                      dplyr::mutate(baseMeanPRIMED_control = MA) %>% 
                      dplyr::mutate(baseMeanNAIVE_foldChangeModerate = ((AM) / (AA))  ) %>% 
                      dplyr::mutate(baseMeanPRIMED_foldChangeModerate = ((MM) / (MA)) ) %>% 
                      dplyr::mutate(baseMeanNAIVE_foldChangeSevere = ((AS) / (AA))  ) %>% 
                      dplyr::mutate(baseMeanPRIMED_foldChangeSevere = ((MS) / (MA)) ) %>% 
                      dplyr::rename(ControlRatioModerate = wgcna.yall_mod) %>% 
                      dplyr::rename(foldChangeRatioModerate = wgcna.xall_mod) %>% 
                      dplyr::rename(ControlRatioSevere = wgcna.yall_sev) %>% 
                      dplyr::rename(foldChangeRatioSevere = wgcna.xall_sev)
write.csv(Day7_Frontload_2,paste("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Day7_FrontloadedGenes.csv"))


D7Frontloaded <- Day7_Frontload_2 %>% filter(ControlRatioModerate > 1) %>% filter(foldChangeRatioModerate < 1)
for (i in 1:nrow(D7Frontloaded)) { 
 x <-  D7Frontloaded[i,]
  x2 <- x  %>% 
    dplyr::select(!c("foldChangeRatioModerate","ControlRatioModerate","foldChangeRatioSevere","ControlRatioSevere","baseMeanNAIVE_control","baseMeanPRIMED_control","baseMeanNAIVE_foldChangeModerate","baseMeanPRIMED_foldChangeModerate" ,"baseMeanNAIVE_foldChangeSevere","baseMeanPRIMED_foldChangeSevere")) %>% 
    gather(variable, value, -c(Gene,Gene_Description)) %>% 
    dplyr::filter(variable %in% c('AA','AM','MA','MM')) %>% 
    dplyr::mutate(Primary_Treatment = substr(variable, 1,1)) %>% 
    dplyr::mutate(Second_Treatment = substr(variable, 2,2)) %>% 
    # dplyr::group_by(Primary_Treatment, Third_Treatment) %>% 
    # dplyr::summarise(meanEXP = mean(value), sdExp = sd(value)) %>% 
      ggplot(aes(x=Second_Treatment , y=value, fill=Primary_Treatment)) +  # , colour=supp, group=supp))
        theme_classic() +
        #geom_errorbar(aes(ymin=meanEXP-sdExp, ymax=meanEXP+sdExp), colour="black", width=.1, position=pd) +
        # geom_point(position=pd, size = 4, shape=21) +   
        geom_bar(position=position_dodge(), aes(y=value), stat="identity", width=0.5) +
        xlab("Second pCO2 treatment") +
        ylab('Gene Expression (averaged raw data)') +                 # note the mean was first by sample ID THEN by treatment
        scale_fill_manual(values=c("grey85","grey50")) +
        #scale_color_manual(values=c("#56B4E9","#E69F00")) +
        # ggtitle(paste("Day 7:",apriori_DESeq2_condenced[i,1],":", apriori_DESeq2_condenced[i,3],sep='')) +
        ggtitle(gsub(" .*", "\\1", D7Frontloaded[i,2])) +
        # expand_limits(y=0) +                                                    # Expand y range
        # scale_y_continuous(limits=c((min_p1), (max_p1))) +
        # scale_y_continuous(limits = c((max(sdExp)+0.5),(min(sdExp)-0.5)))  +
        # ylim(ExpMin, ExpMin) +
        theme(axis.text.x = element_text(size = 20),
              axis.text.y = element_text(size = 20),
              axis.ticks.length=unit(.25, "cm"))+
        theme(legend.position = "none")
  
  gene <- D7Frontloaded[i,1]
  title <- gsub(" .*", "\\1", (sub(" ", "_",  D7Frontloaded[i,2])) )
  
pdf(paste0("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Frontloading_plots/Day7/",gene,"_", title,".pdf", sep = ''))
print(x2)
dev.off()

  }














# Day 14 gene expression
Day_WGCNA_genes <- Day14_exp_data %>% 
  dplyr::filter(Gene %in% NaiveResponse_genes)
Day14_WGCNA_genes_melted     <- Day_WGCNA_genes %>% 
  reshape2::melt(id.var = 'Gene') %>% 
  dplyr::rename(Sample.Name = variable)
Day14_WGCNA_genes_Merge <- merge(Day14_WGCNA_genes_melted, Day14_Master.Exp.Metadata, by = 'Sample.Name') %>% 
  dplyr::group_by(Gene, All_Treatment) %>% 
  dplyr::select(!'Sample.Name') %>% 
  dplyr::summarise(meanExp = mean(value))
Day14_READY <- dcast(Day14_WGCNA_genes_Merge, Gene ~ All_Treatment)
for (i in 1:nrow(Day14_READY)) {
          # Moderate - higher expression AM > AA
        if (Day14_READY$AM[i] > Day14_READY$AA[i]) {
          Day14_READY$wgcna.xall_mod[i] <- ( (Day14_READY$MM[i] / Day14_READY$MA[i]) / (Day14_READY$AM[i] / Day14_READY$AA[i]) ) # call MA as the control for MM ratio
          Day14_READY$wgcna.yall_mod[i] <- (Day14_READY$MA[i] / Day14_READY$AA[i]) # Y Axis - this is simply the conditioned control over the naive control ( MA / AA ) 
              } else {
                Day14_READY$wgcna.xall_mod[i] <- NA  # X axis  - call NA    
                Day14_READY$wgcna.yall_mod[i] <- NA # Y Axis  - call NA
              }
            # Severe - higher expression AS > AA
          if (Day14_READY$AS[i] > Day14_READY$AA[i]) {
            Day14_READY$wgcna.xall_sev[i] <- ( (Day14_READY$MS[i] / Day14_READY$MA[i]) / (Day14_READY$AS[i] / Day14_READY$AA[i]) ) # call AA as the control for MM ratio
            Day14_READY$wgcna.yall_sev[i] <- (Day14_READY$MA[i] / Day14_READY$AA[i])            # Y Axis - this is simply the conditioned control over the naive control ( MA / AA ) 
                } else {
                  Day14_READY$wgcna.xall_sev[i] <- NA  # X axis  - call NA 
                  Day14_READY$wgcna.yall_sev[i] <- NA # Y Axis  - call NA
                }
}
# Day14_READY
x = data.frame(Gene = annot$V1[(match(Day14_READY$Gene, annot$V1))],
               Gene_Description = annot$V7[(match(Day14_READY$Gene, annot$V1))])
Day14_Frontload   <- merge(x, Day14_READY, by = 'Gene')
Day14_Frontload_2 <- Day14_Frontload %>% 
                      dplyr::mutate(baseMeanNAIVE_control = AA) %>% 
                      dplyr::mutate(baseMeanPRIMED_control = MA) %>% 
                      dplyr::mutate(baseMeanNAIVE_foldChangeModerate = ((AM) / (AA))  ) %>% 
                      dplyr::mutate(baseMeanPRIMED_foldChangeModerate = ((MM) / (MA)) ) %>% 
                      dplyr::mutate(baseMeanNAIVE_foldChangeSevere = ((AS) / (AA))  ) %>% 
                      dplyr::mutate(baseMeanPRIMED_foldChangeSevere = ((MS) / (MA)) ) %>% 
                      dplyr::rename(ControlRatioModerate = wgcna.yall_mod) %>% 
                      dplyr::rename(foldChangeRatioModerate = wgcna.xall_mod) %>% 
                      dplyr::rename(ControlRatioSevere = wgcna.yall_sev) %>% 
                      dplyr::rename(foldChangeRatioSevere = wgcna.xall_sev)
write.csv(Day14_Frontload_2,paste("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Day14_FrontloadedGenes.csv"))


D14Frontloaded <- Day14_Frontload_2 %>% filter(ControlRatioModerate > 1) %>% filter(foldChangeRatioModerate < 1)
for (i in 1:nrow(D14Frontloaded)) { 
 x <-  D14Frontloaded[i,]
  x2 <- x  %>% 
    dplyr::select(!c("foldChangeRatioModerate","ControlRatioModerate","foldChangeRatioSevere","ControlRatioSevere","baseMeanNAIVE_control","baseMeanPRIMED_control","baseMeanNAIVE_foldChangeModerate","baseMeanPRIMED_foldChangeModerate" ,"baseMeanNAIVE_foldChangeSevere","baseMeanPRIMED_foldChangeSevere")) %>% 
    gather(variable, value, -c(Gene,Gene_Description)) %>% 
    dplyr::filter(variable %in% c('AA','AM','MA','MM')) %>% 
    dplyr::mutate(Primary_Treatment = substr(variable, 1,1)) %>% 
    dplyr::mutate(Second_Treatment = substr(variable, 2,2)) %>% 
    # dplyr::group_by(Primary_Treatment, Third_Treatment) %>% 
    # dplyr::summarise(meanEXP = mean(value), sdExp = sd(value)) %>% 
      ggplot(aes(x=Second_Treatment , y=value, fill=Primary_Treatment)) +  # , colour=supp, group=supp))
        theme_classic() +
        #geom_errorbar(aes(ymin=meanEXP-sdExp, ymax=meanEXP+sdExp), colour="black", width=.1, position=pd) +
        # geom_point(position=pd, size = 4, shape=21) +   
        geom_bar(position=position_dodge(), aes(y=value), stat="identity", width=0.5) +
        xlab("Second pCO2 treatment") +
        ylab('Gene Expression (averaged raw data)') +                 # note the mean was first by sample ID THEN by treatment
        scale_fill_manual(values=c("grey85","grey50")) +
        #scale_color_manual(values=c("#56B4E9","#E69F00")) +
        # ggtitle(paste("Day 7:",apriori_DESeq2_condenced[i,1],":", apriori_DESeq2_condenced[i,3],sep='')) +
        ggtitle(gsub(" .*", "\\1", D14Frontloaded[i,2])) +
        # expand_limits(y=0) +                                                    # Expand y range
        # scale_y_continuous(limits=c((min_p1), (max_p1))) +
        # scale_y_continuous(limits = c((max(sdExp)+0.5),(min(sdExp)-0.5)))  +
        # ylim(ExpMin, ExpMin) +
        theme(axis.text.x = element_text(size = 20),
              axis.text.y = element_text(size = 20),
              axis.ticks.length=unit(.25, "cm"))+
        theme(legend.position = "none")

  gene <- D14Frontloaded[i,1]
  title <- gsub(" .*", "\\1", (sub(" ", "_",  D14Frontloaded[i,2])) )
  
pdf(paste0("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Frontloading_plots/Day14/",gene, "_", title,".pdf", sep = ''))
print(x2)
dev.off()

  }











# Day 21 gene expression
Day_WGCNA_genes <- Day21_exp_data %>% 
  dplyr::filter(Gene %in% NaiveResponse_genes)
Day21_WGCNA_genes_melted     <- Day_WGCNA_genes %>% 
  reshape2::melt(id.var = 'Gene') %>% 
  dplyr::rename(Sample.Name = variable)
Day21_WGCNA_genes_Merge <- merge(Day21_WGCNA_genes_melted, Day21_Master.Exp.Metadata, by = 'Sample.Name') %>% 
  dplyr::group_by(Gene, All_Treatment) %>% 
  dplyr::select(!'Sample.Name') %>% 
  dplyr::summarise(meanExp = mean(value))
Day21_READY <- dcast(Day21_WGCNA_genes_Merge, Gene ~ All_Treatment)
for (i in 1:nrow(Day21_READY)) {
          # Moderate - higher expression AM > AA
        if (Day21_READY$AAM[i] > Day21_READY$AAA[i]) {
          Day21_READY$wgcna.xall_mod[i] <- ( (Day21_READY$MAM[i] / Day21_READY$MAA[i]) / (Day21_READY$AAM[i] / Day21_READY$AAA[i]) ) # call MA as the control for MM ratio
          Day21_READY$wgcna.yall_mod[i] <- (Day21_READY$MAA[i] / Day21_READY$AAA[i]) # Y Axis - this is simply the conditioned control over the naive control ( MA / AA ) 
              } else {
                Day21_READY$wgcna.xall_mod[i] <- NA  # X axis  - call NA    
                Day21_READY$wgcna.yall_mod[i] <- NA # Y Axis  - call NA
              }
}
# Day21_READY


x = data.frame(Gene = annot$V1[(match(Day21_READY$Gene, annot$V1))],
               Gene_Description = annot$V7[(match(Day21_READY$Gene, annot$V1))])
Day21_Frontload   <- merge(x, Day21_READY, by = 'Gene')
Day21_Frontload_2 <- Day21_Frontload %>% 
                      dplyr::mutate(baseMeanNAIVE_control = AAA) %>% 
                      dplyr::mutate(baseMeanPRIMED_control = MAA) %>% 
                      dplyr::mutate(baseMeanNAIVE_foldChangeModerate = ((AAM) / (AAA))  ) %>% 
                      dplyr::mutate(baseMeanPRIMED_foldChangeModerate = ((MAM) / (MAA)) ) %>% 
                      dplyr::rename(ControlRatio = wgcna.yall_mod) %>% 
                      dplyr::rename(foldChangeRatio = wgcna.xall_mod)
write.csv(Day21_Frontload_2,paste("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Day21_FrontloadedGenes.csv"))


library(tidyverse)
D21Frontloaded <- Day21_Frontload_2 %>% filter(ControlRatio > 1) %>% filter(foldChangeRatio < 1)
for (i in 1:nrow(D21Frontloaded)) { 
 x <-  D21Frontloaded[i,]
  x2 <- x  %>% 
    dplyr::select(!c('foldChangeRatio','ControlRatio','baseMeanNAIVE_control','baseMeanPRIMED_control','baseMeanNAIVE_foldChangeModerate','baseMeanPRIMED_foldChangeModerate')) %>% 
    gather(variable, value, -c(Gene,Gene_Description)) %>% 
    dplyr::filter(variable %in% c('AAA','AAM','MAA','MAM')) %>% 
    dplyr::mutate(Primary_Treatment = substr(variable, 1,1)) %>% 
    dplyr::mutate(Second_Treatment = substr(variable, 2,2)) %>% 
    dplyr::mutate(Third_Treatment = substr(variable, 3,3)) %>% 
    # dplyr::group_by(Primary_Treatment, Third_Treatment) %>% 
    # dplyr::summarise(meanEXP = mean(value), sdExp = sd(value)) %>% 
      ggplot(aes(x=Third_Treatment , y=value, fill=Primary_Treatment)) +  # , colour=supp, group=supp))
        theme_classic() +
        #geom_errorbar(aes(ymin=meanEXP-sdExp, ymax=meanEXP+sdExp), colour="black", width=.1, position=pd) +
        # geom_point(position=pd, size = 4, shape=21) +   
        geom_bar(position=position_dodge(), aes(y=value), stat="identity", width=0.5) +
        xlab("Third pCO2 treatment") +
        ylab('Gene Expression (averaged raw data)') +                 # note the mean was first by sample ID THEN by treatment
        scale_fill_manual(values=c("grey85","grey50")) +
        #scale_color_manual(values=c("#56B4E9","#E69F00")) +
        # ggtitle(paste("Day 7:",apriori_DESeq2_condenced[i,1],":", apriori_DESeq2_condenced[i,3],sep='')) +
        ggtitle(gsub(" .*", "\\1", D21Frontloaded[i,2])) +
        # expand_limits(y=0) +                                                    # Expand y range
        # scale_y_continuous(limits=c((min_p1), (max_p1))) +
        # scale_y_continuous(limits = c((max(sdExp)+0.5),(min(sdExp)-0.5)))  +
        # ylim(ExpMin, ExpMin) +
        theme(axis.text.x = element_text(size = 20),
              axis.text.y = element_text(size = 20),
              axis.ticks.length=unit(.25, "cm"))+
        theme(legend.position = "none")
  
  gene <- D21Frontloaded[i,1]
  title <- gsub(" .*", "\\1", (sub(" ", "_",  D21Frontloaded[i,2])) )
  
pdf(paste0("C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/Analysis/Output/WGCNA/subseq_treatments_all/Frontloading_plots/Day21/",gene, "_", title,".pdf", sep = ''))
print(x2)
dev.off()

  }
```

### Day 7 - Frontloaded genes (Plot)

```{r plotting Day 7, echo=FALSE}
# plot the frontloaded genes formula using x axis for the moderate treatment (second exposure, day 7)
P.wgcna.mod <- Day7_READY %>% 
  dplyr::select(c('wgcna.xall_mod', 'wgcna.yall_mod')) %>% 
  na.omit() %>% 
  dplyr::filter(wgcna.yall_mod < 3) %>% 
  ggplot(aes(x=wgcna.xall_mod, y=wgcna.yall_mod)) +
  geom_point() +
  theme_classic() + 
  stat_smooth(method = "lm", 
              formula = y ~ x + poly(x, 2) - 1) +
  geom_vline(xintercept=1, linetype="dotted") + 
  geom_hline(yintercept=1, linetype="dotted") + 
  labs(y= "Conditioned to naive control ratio", 
       x = "Conditioned to naive foldchange ratio",
       title = "Day 7 
       Frontloaded genes; response to moderate pCO2") + 
  expand_limits(x = 0, y = 0) + 
  annotate("rect", xmin = 0, xmax = 1, ymin = 1, ymax = 1.5,
           alpha = .2)
P.wgcna.mod


# plot the frontloaded genes formula using x axis for the severe treatment (second exposure, day 7)
P.wgcna.sev <- Day7_READY %>% 
  dplyr::select(c('wgcna.xall_sev', 'wgcna.yall_sev')) %>% 
  na.omit() %>% 
  dplyr::filter(wgcna.yall_sev < 3) %>% 
  ggplot(aes(x=wgcna.xall_sev, y=wgcna.yall_sev)) +
  geom_point() +
  theme_classic() + 
  stat_smooth(method = "lm", 
              formula = y ~ x + poly(x, 2) - 1) +
  geom_vline(xintercept=1, linetype="dotted") + 
  geom_hline(yintercept=1, linetype="dotted") + 
  labs(y= "Conditioned to naive control ratio", 
       x = "Conditioned to naive foldchange ratio",
       title = "Day 7 Frontloaded genes; response to severe pCO2") + 
  expand_limits(x = 0, y = 0) + 
  annotate("rect", xmin = 0, xmax = 1, ymin = 1, ymax = 1.5,
           alpha = .2) 
P.wgcna.sev

```

### Day 7 - Frontloaded genes (Table)

```{r table Day 7, echo=TRUE}

# MODERATE second treatment
Day7.wgcna.frontloaded_genes_mod <- Day7_READY %>% 
                                dplyr::filter(wgcna.xall_mod < 1) %>% 
                                dplyr::filter(wgcna.yall_mod > 1) %>%
                                dplyr::select('Gene')

Day7.wgcna.frontloadprobes_mod        = Day7.wgcna.frontloaded_genes_mod$Gene
probes2annot_mod                      = match(Day7.wgcna.frontloadprobes_mod, annot$V1)
Day7.wgcna.frontloaded_mod_ANNOT      = data.frame(geneSymbol = annot$V1[probes2annot_mod],
                                              Genes = annot$V7[probes2annot_mod])
kable(Day7.wgcna.frontloaded_mod_ANNOT)

# SEVERE second treatment
Day7.wgcna.frontloaded_genes_sev <- Day7_READY %>% 
                                dplyr::filter(wgcna.xall_sev < 1) %>% 
                                dplyr::filter(wgcna.yall_sev > 1) %>%
                                dplyr::select('Gene')

Day7.wgcna.frontloadprobes_sev        = Day7.wgcna.frontloaded_genes_sev$Gene
probes2annot_sev                      = match(Day7.wgcna.frontloadprobes_sev, annot$V1)
Day7.wgcna.frontloaded_sev_ANNOT      = data.frame(geneSymbol = annot$V1[probes2annot_sev],
                                              Genes = annot$V7[probes2annot_sev])
kable(Day7.wgcna.frontloaded_sev_ANNOT)
```

### Day 14 - Frontloaded genes (Plot)

```{r plotting Day 14, echo=F}
# plot the frontloaded genes formula using x axis for the moderate treatment (second exposure, day 14)
P.wgcna.mod <- Day14_READY %>% 
  dplyr::select(c('wgcna.xall_mod', 'wgcna.yall_mod')) %>% 
  na.omit() %>% 
  #dplyr::filter(wgcna.yall_mod < 3) %>% 
  ggplot(aes(x=wgcna.xall_mod, y=wgcna.yall_mod)) +
  geom_point() +
  theme_classic() + 
  stat_smooth(method = "lm", 
              formula = y ~ x + poly(x, 2) - 1) +
  geom_vline(xintercept=1, linetype="dotted") + 
  geom_hline(yintercept=1, linetype="dotted") + 
  labs(y= "Conditioned to naive control ratio", 
       x = "Conditioned to naive foldchange ratio",
       title = "Day 14 
       Frontloaded genes; response to moderate pCO2") + 
  expand_limits(x = 0, y = 0) + 
  annotate("rect", xmin = 0, xmax = 1, ymin = 1, ymax = 1.5,
           alpha = .2)
P.wgcna.mod


# plot the frontloaded genes formula using x axis for the severe treatment (second exposure, day 14)
P.wgcna.sev <- Day14_READY %>% 
  dplyr::select(c('wgcna.xall_sev', 'wgcna.yall_sev')) %>% 
  dplyr::filter(wgcna.xall_sev < 4) %>% 
  na.omit() %>% 
  #dplyr::filter(wgcna.yall_sev < 3) %>% 
  ggplot(aes(x=wgcna.xall_sev, y=wgcna.yall_sev)) +
  geom_point() +
  theme_classic() + 
  stat_smooth(method = "lm", 
              formula = y ~ x + poly(x, 2) - 1) +
  geom_vline(xintercept=1, linetype="dotted") + 
  geom_hline(yintercept=1, linetype="dotted") + 
  labs(y= "Conditioned to naive control ratio", 
       x = "Conditioned to naive foldchange ratio",
       title = "Day 14 Frontloaded genes; response to severe pCO2") + 
  expand_limits(x = 0, y = 0) + 
  annotate("rect", xmin = 0, xmax = 1, ymin = 1, ymax = 1.5,
           alpha = .2) 
P.wgcna.sev

```

### Day 14 - Frontloaded genes (Table)

```{r table Day 14, echo=T}
# MODERATE second treatment
Day14.wgcna.frontloaded_genes_mod <- Day14_READY %>% 
                                dplyr::filter(wgcna.xall_mod < 1) %>% 
                                dplyr::filter(wgcna.yall_mod > 1) %>%
                                dplyr::select('Gene') # %>% 

Day14.wgcna.frontloadprobes_mod        = Day14.wgcna.frontloaded_genes_mod$Gene
probes2annot_mod                       = match(Day14.wgcna.frontloadprobes_mod, annot$V1)
Day14.wgcna.frontloaded_mod_ANNOT      = data.frame(geneSymbol = annot$V1[probes2annot_mod],
                                         Genes = annot$V7[probes2annot_mod])
kable(Day14.wgcna.frontloaded_mod_ANNOT)

# SEVERE second treatment
Day14.wgcna.frontloaded_genes_sev <- Day14_READY %>% 
                                dplyr::filter(wgcna.xall_sev < 1) %>% 
                                dplyr::filter(wgcna.yall_sev > 1) %>%
                                dplyr::select('Gene') # %>% 

Day14.wgcna.frontloadprobes_sev        = Day14.wgcna.frontloaded_genes_sev$Gene
probes2annot_sev                       = match(Day14.wgcna.frontloadprobes_sev, annot$V1)
Day14.wgcna.frontloaded_sev_ANNOT      = data.frame(geneSymbol = annot$V1[probes2annot_sev],
                                              Genes = annot$V7[probes2annot_sev])
kable(Day14.wgcna.frontloaded_sev_ANNOT)
```

### Day 21 - Frontloaded genes (Plot)

```{r plotting Day 21, echo=F}
# plot the frontloaded genes formula using x axis for the moderate treatment (second exposure, day 21)
par(mfrow=c(1,2))
P.wgcna.mod <- Day21_READY %>% 
  dplyr::select(c('wgcna.xall_mod', 'wgcna.yall_mod')) %>% 
  na.omit() %>% 
  dplyr::filter(wgcna.yall_mod < 4) %>% 
  ggplot(aes(x=wgcna.xall_mod, y=wgcna.yall_mod)) +
  geom_point() +
  theme_classic() + 
  stat_smooth(method = "lm", 
              formula = y ~ x + poly(x, 2) - 1) +
  geom_vline(xintercept=1, linetype="dotted") + 
  geom_hline(yintercept=1, linetype="dotted") + 
  labs(y= "Conditioned to naive control ratio", 
       x = "Conditioned to naive foldchange ratio",
       title = "Day 21 
       Frontloaded genes; response to moderate pCO2") + 
  expand_limits(x = 0, y = 0) + 
  annotate("rect", xmin = 0, xmax = 1, ymin = 1, ymax = 4.5,
           alpha = .2) 
P.wgcna.mod

```

### Day 21 - Frontloaded genes (Table)

```{r table Day 21, echo=T}
# MODERATE second treatment
Day21.wgcna.frontloaded_genes_mod <- Day21_READY %>% 
                                dplyr::filter(wgcna.xall_mod < 1) %>% 
                                dplyr::filter(wgcna.yall_mod > 1) %>%
                                dplyr::select('Gene') # %>% 

Day21.wgcna.frontloadprobes_mod        = Day21.wgcna.frontloaded_genes_mod$Gene
probes2annot_mod                       = match(Day21.wgcna.frontloadprobes_mod, annot$V1)
Day21.wgcna.frontloaded_mod_ANNOT      = data.frame(geneSymbol = annot$V1[probes2annot_mod],
                                         Genes = annot$V7[probes2annot_mod])
kable(Day21.wgcna.frontloaded_mod_ANNOT)
```

```{r summary table, echo = T} 


View(rbind(Day7.wgcna.frontloaded_mod_ANNOT, Day14.wgcna.frontloaded_mod_ANNOT, Day21.wgcna.frontloaded_mod_ANNOT) %>% 
  dplyr::count(Genes))

```