---
title: "WGCNA_data_prep"
author: "Samuel Gurr"
date: "February 4, 2021"
output: html_document
---

# README: Objective of this script is to merge the following data.frames/tables
(1) 'Master_Exp.Treatment_Metadata.csv' - output from Count.Matrix_Stats.Filter.Rmd. This file contains all treatment data, expoeriment date, replicate tank ID, and sample name 
(2) Phenotype data from Gurr et al. 2020 (bioRxiv, in review in JEB) - all data was loaded into the WGCNA folder. Data from github
NOTE: TagSeq gene exoression data is NOT from the same animals with phenotypic measurements BUT from the sample replicate tank and timepoint.
a.) "Gurr.etal_TAOC_Protein_Organic.biomass" = total antioxidant capacity, total protein, and organic biomass AFDW taken from on the same date and replicate tank
b.) "Gurr.etal_Shell.length"                 = shell legnth of animals used for respirometry (N = 3 per replicate tank-1 timepoint-1)
c.) "Gurr.etal_Respiration.Biovolume"        = respiration rate (ug-1 hour-1 individual-1) and biovoume measurements (N = 1 per replicate tank-1 timepoint-1; measured as the mean for each PreSens respiration vial)

# Set-up

### Load libraries
```{r  load library}
library(dplyr)
library(reshape2)
print(getwd()) # working directory is the scipts folder 
```

# LOAD DATA:

```{r  load_data}
print(getwd())  # working directory is the scipts folder 
Master.treatment.data <- read.csv("../../Analysis/Data/Master_Exp.Treatment_Metadata.csv", sep=',', header=TRUE) # read the output count matrix - NOTE: TRIMMED at 30 phred threshold!
Resp.Biovol.master       <- read.csv(file="../../Analysis/Data/WGCNA/Gurr.etal_Respiration.Biovolume.csv", sep=',', header=TRUE)
Shell.Length.master      <- read.csv(file="../../Analysis/Data/WGCNA/Gurr.etal_Shell.length.csv", sep=',', header=TRUE)
TAOC.Protein.AFDW.master <- read.csv(file="../../Analysis/Data/WGCNA/Gurr.etal_TAOC_Protein_Organic.biomass.csv", sep=',', header=TRUE)
```

### summarize and merge data by tank ID
```{r  Shell.Length}
unique(Shell.Length.master$Date)
Shell.Length <- Shell.Length.master %>% 
  dplyr::filter(Date %in% c(20190731,20190807, 20190814)) %>%  # call   "DAY7"  "DAY14" "DAY21"
   dplyr::group_by(Date,Tank.ID) %>% 
  summarise(mean.Shell.Length = mean(Length))
Shell.Length <- na.omit(Shell.Length)
dim(Shell.Length)
Shell.Length$Date_Tank.ID <- paste(Shell.Length$Date, Shell.Length$Tank.ID, sep= "_")
Shell.Length <- melt(Shell.Length, id.vars = c('Date','Tank.ID','Date_Tank.ID'))
```


```{r  Resp.Biovol}
unique(Resp.Biovol.master$Date)
Resp.Biovol <- Resp.Biovol.master %>% 
  dplyr::select(c(1,2,3,15,16)) %>% 
  dplyr::filter(Date %in% c(20190731,20190807, 20190814)) %>%  # call "DAY7"  "DAY14" "DAY21"
  dplyr::group_by(Date,Tank.ID) %>% 
  dplyr::summarise(mean.Resp.ugLhrindiv = mean(resp.COUNT.µg.L.hr.indiv), # note resp rate per individual - corrected from raw data on 3 individuals measurement-1
                                              mean.Biovol = mean(mean_biovolume))
dim(Resp.Biovol)
Resp.Biovol$Date_Tank.ID <- paste(Resp.Biovol$Date, Resp.Biovol$Tank.ID, sep= "_")
Resp.Biovol <- melt(Resp.Biovol, id.vars = c('Date','Tank.ID','Date_Tank.ID'))
```

```{r  TAOC.Protein.AFDW}

TAOC.Protein.AFDW.master.OM <- TAOC.Protein.AFDW.master %>%  dplyr::filter(!ID %in% c('1477', '1438', '1594', '1570', '1606', '1551', '1563','349', '357', '337',  '450','1588', '1483'))
# ommit the duplicates that were either negative TAOC or rerun due to tech error
TAOC.Protein.AFDW <- TAOC.Protein.AFDW.master.OM %>% 
  dplyr::select(c(1,3:8)) %>% 
  dplyr::filter(Date %in% c(20190731,20190814)) %>%  # call   "DAY7"  "DAY14" "DAY21"
  dplyr::group_by(Date, Tank.ID) %>% 
  dplyr::summarise(mean.µmol.CRE.g.protein = mean(µmol.CRE.g.protein), 
                   mean.AFDW = mean(mgTOTAL_AFDW), 
                   mean.mgProt.mgAFDW = mean(mgProtein.mgAFDW))
dim(TAOC.Protein.AFDW)
TAOC.Protein.AFDW$Date_Tank.ID <- paste(TAOC.Protein.AFDW$Date, TAOC.Protein.AFDW$Tank.ID, sep= "_")
TAOC.Protein.AFDW <- melt(TAOC.Protein.AFDW, id.vars = c('Date','Tank.ID','Date_Tank.ID'))
```

```{r  rbind_phys}

Master.Phys.data <- rbind(TAOC.Protein.AFDW, Resp.Biovol, Shell.Length)
Master.Phys.data<- Master.Phys.data[-3]
dim(Master.Phys.data)
```

```{r  rbind_treament.data}

Treatment_Date <- Master.treatment.data %>%  
  dplyr::filter(Date %in% c(20190731,20190807,20190814)) %>% 
  dplyr::select('Date','Tank.ID', 'All_Treatment', 'Primary_Treatment', 'Second_Treament' ,'Third_Treatment') 
Treatment_Date.melt <- melt(Treatment_Date, id.vars = c('Date','Tank.ID'))

Master.Metadata <- rbind(Master.Phys.data, Treatment_Date.melt)
```



```{r  merge_TagSeqSampleID}
Seq.ID_Date <- Master.treatment.data %>% dplyr::select('Sample.Name','Date','Tank.ID')

# common column to bind
Seq.ID_Date$Date_Tank.ID <- paste(Seq.ID_Date$Date,Seq.ID_Date$Tank.ID)
Master.Metadata$Date_Tank.ID <- paste(Master.Metadata$Date,Master.Metadata$Tank.ID)

Master.Metadata_TagSeq.ID <- merge(Master.Metadata, Seq.ID_Date, by = 'Date_Tank.ID')

Master.Metadata_TagSeq.ID.dcast <- dcast(Master.Metadata_TagSeq.ID, Sample.Name ~ variable)

Master.Metadata_TagSeq.ID.dcast.merge <-  merge(Master.Metadata_TagSeq.ID.dcast, Seq.ID_Date[1:3], by = 'Sample.Name')

colnames(Master.Metadata_TagSeq.ID.dcast.merge)

Master.FINAL <-  Master.Metadata_TagSeq.ID.dcast.merge[,c(12,13,1,8:11,2:7)] %>% group_by(Date) %>% arrange(desc(Primary_Treatment), .by_group = TRUE)
```

```{r  write_output}
write.csv(Master.FINAL,paste(path.out,"Master_Phyenotype.and.Exp.Treatment_Metadata.csv"))
```




